---
title: "Visualize Data"
---

## Your Turn 0

Add a setup chunk that loads the tidyverse packages.



## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}

```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

## Help Me

What do `facet_grid()` and `facet_wrap()` do? 
(run the code and interpret)

```{r}
# Makes a plot that the commands below will modify
q <- ggplot(mpg) + geom_point(aes(x = displ, y = hwy))

q + facet_grid(. ~ cyl)
q + facet_grid(drv ~ .)
q + facet_grid(drv ~ cyl)
q + facet_wrap(~ class)
```


## Your Turn 3

Replace this scatterplot with one that draws boxplots. Use the cheatsheet or try your best guess.

```{r}
ggplot(mpg) + geom_point(aes(class, hwy))
```

## Your Turn 4

Make a histogram of the `hwy` variable from dataset `mpg`. Hint: do not supply a y variable.

```{r}

```

## Your Turn 5

Use the help page for `geom_histogram` to make the bins 2 units wide.

```{r}

```

## Your Turn 6

Make a density chart of `hwy` colored by `class`. Use the help page for `geom_density` to choose a "color" aesthetic for class.

```{r}

```

## Your Turn 7

What will this code do? Private chat with your class partner.

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))

# you can run the code directly to see what it does. BUT,
# before you do that, just read the code and predict in your
# head what this code will do.
```

## Quiz

What is different about this plot? Run the code!

```{r}
p <- ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))

library(plotly)
ggplotly(p)
```

Detach package `plotly`. `plotly` has many functions that have same names as functions in the package `dplyr`, and to avoid any conflict between functions, we are unloading package `plotly` for now.

```{r}
detach("package:plotly", unload = TRUE)
```

## Your Turn 8

What does `getwd()` return?

```{r}

```

## Your Turn 9

Save your last plot and then locate it in your files list. You may have to refresh your file listing.

```{r}

```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

# Bonus code (run in your own time)

## Heatmaps

There are several methods in R to create a heatmap. Since, I have introduced you to **ggplot2** already, lets explore how you can make one using one of the <geom_function>. More heatmap code can be found at https://www.r-graph-gallery.com/79-levelplot-with-ggplot2.html

```{r}
# import data
# this data is from USDA Ag Data Commons
# https://data.nal.usda.gov/
alfalfa <- readxl::read_xlsx("ALFALFA-NAL.xlsx", sheet = "DATA") %>% 
  group_by(Year, State) %>% 
  summarize(yield = mean(MYield))

# look at first few rows of dataframe
head(alfalfa)

# create heatmap using `geom-tile`
heatmap_1 <- ggplot(data = alfalfa,
                    mapping = aes(x = Year,
                                  y = State,
                                  fill = yield)) +
  geom_tile() # this is the geom_function for creating heatmaps

heatmap_1

# make the heatmap interactive
library(plotly)
ggplotly(heatmap_1)
```

## 3D maps

The package `rayshader` has the ability to turn your **ggplot2** objects into 3D maps.

```{r}
# create a 2D plot
mtplot <- ggplot(mtcars) + 
  geom_point(aes(x = mpg, y = disp, color = cyl)) + 
  scale_color_continuous(limits = c(0, 8))

mtplot

# convert it to 3D
library(rayshader)
plot_gg(mtplot) # this code may not work in RStudio Cloud, but it works on 
                # RStudio Desktop. If it does not, please let me know. To
                # learn more about rayshader, please go to
                # https://www.rayshader.com/

# detach plotly
detach("package:plotly", unload = TRUE)
```


