---
title: "Transform Data"
---

```{r setup}
library(tidyverse)

# Toy datasets to use

pollution <- tribble(
       ~city,   ~size, ~amount, 
  "New York", "large",      23,
  "New York", "small",      14,
    "London", "large",      22,
    "London", "small",      16,
   "Beijing", "large",      121,
   "Beijing", "small",      56
)

# library(babynames)
```

# Import Data

## Your Turn 0

Set the current working directory(you can use the function `setwd()`) to "/cloud/project/03-Transform". Then, import the `babynames.csv` data set, which is in the `03-Transform` folder, as `babynames`. Then copy the import code into the code chunk below. Does it run?

What will you change in the code to make it run?

```{r}

```

```{r}
babynames
```

# dplyr

## Your Turn 1

Run the code, and see what it does. After that, alter the code to select just the `n` column:

```{r}
select(babynames, name, prop)
```

## Quiz

Which of these is NOT a way to select the `name` and `n` columns together?

```{r}
select(babynames, -c(year, sex, prop))
select(babynames, name:n)
select(babynames, starts_with("n"))
select(babynames, ends_with("n"))
```

## Your Turn 2

Use `filter`, `babynames`, and the logical operators to find:

* All of the names where prop is greater than or equal to 0.08  
* All of the children named "Sea"  
* All of the baby names from year 2016

Note: the function `filter()` is found in multiple packages. When you start typing `filter` in the chunk below, you may see the first option as `filter{plotly}` i.e. this is the function `filter()` from `plotly` package. We don't want to use this one here. We are looking for one from the `dplyr` package. One way to do that is to put the package name before the function like this `dplyr::filter`. or we can detach the `plotly` package.

```{r}

```

## Your Turn 3

Use Boolean operators to alter the code below to return only the rows that contain:

* _Girls_ named Lucky  
* Names that were used by exactly 5 or 6 children in 1880  
* Names that are one of Acura, Lexus, or Yugo

```{r}

```

## Your Turn 4

Arrange `babynames` by n. Add prop as a second (tie breaking) variable to arrange on.

Can you tell the smallest value of n?
What is the largest value of n?

```{r}

```

## Your Turn 5

Use `desc()` to find the names with the highest prop.

Then, use `desc()` to find the names with the highest n.

```{r}

```

## Your Turn 6

Use `%>%` to write a sequence of functions that: 

1. Filters babynames to just the girls that were born in 2017, *then...*  
2. Selects the `name` and `n` columns, *then...*  
3. Arranges the results so that the most popular names are near the top.

```{r}

```

## Your Turn 7

1. Trim `babynames` to just the rows that contain your `name` and your `sex`  
2. Trim the result to just the columns that will appear in your graph (not strictly necessary, but useful practice)  
3. Plot the results as a line graph with `year` on the x axis and `prop` on the y axis

```{r}

```

## Your Turn 8

Use `summarise()` to compute three statistics about the data:

1. The first (minimum) year in the dataset  
2. The last (maximum) year in the dataset  
3. The total number of children represented in the data

```{r}

```

## Your Turn 9

Copmplete the code below to extract the rows where `name == "Khaleesi"`. Then use `summarise()` and `sum()` and `min()` to find:

1. The total number of children named Khaleesi
2. The first year Khaleesi appeared in the data

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
babynames ___ 
  filter(_______________________) ___
  ___________(total = ________, first = _______)
```


## Your Turn 10

Use `group_by()`, `summarise()`, and `arrange()` to display the ten most popular names. Compute popularity as the *total* number of children of a single gender given a name.

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
babynames %>%
  _______(name, sex) %>% 
  _______(total = _____(n)) %>% 
  _______(desc(_____))
```


## Your Turn 11

Use `group_by()` to calculate and then plot the total number of children born each year over time.

```{r}

```

## Your Turn 12

Fill in the blanks to: 
Rank each name _within its year and sex_. 

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
babynames %>% 
  group_by(____, ____) %>% 
  ______(rank = min_rank(____(prop)))
```

## Your Turn 13

Fill in the blanks to: 
Compute the median of the ranks for each name (i.e., _for each combination of name and sex_). Arrange the results from lowest median rank to highest

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
babynames %>% 
  group_by(year, sex) %>% 
  mutate(rank = min_rank(desc(prop))) %>% 
  group_by(____, ____) %>% 
  ________(median_rank = ______(_____)) %>% 
  arrange(median_rank)

```

***

# Take aways

* Extract variables with `select()`  
* Extract cases with `filter()`  
* Arrange cases, with `arrange()`  

* Make tables of summaries with `summarise()`  
* Make new variables, with `mutate()`  
* Do groupwise operations with `group_by()`

* Connect operations with `%>%`  


